Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-09-05 apj.tex
making way through Greg comment

Each supernova is described by its parameters
 $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, $E_{\gamma^1}(B-V)$,  $A_{\delta U}=\delta_U D$.
 Their Pearson correlation coefficients are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V),  A_{\delta U}) =\\
\begin{pmatrix}
\begin{array}{rrrrrrr}
1 & 0.01^{+0.06}_{-0.06} & -0.07^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.07} & 0.10^{+0.07}_{-0.06} & 0.00^{+0.08}_{-0.07} & -0.01^{+0.08}_{-0.08} \\
0.01^{+0.06}_{-0.06} & 1 & 0.10^{+0.04}_{-0.04} & -0.26^{+0.03}_{-0.03} & -0.04^{+0.06}_{-0.06} & -0.00^{+0.07}_{-0.08} & -0.00^{+0.08}_{-0.08} \\
-0.07^{+0.07}_{-0.06} & 0.10^{+0.04}_{-0.04} & 1 & -0.14^{+0.03}_{-0.03} & -0.16^{+0.05}_{-0.05} & -0.03^{+0.08}_{-0.08} & 0.01^{+0.08}_{-0.08} \\
-0.04^{+0.06}_{-0.07} & -0.26^{+0.03}_{-0.03} & -0.14^{+0.03}_{-0.03} & 1 & 0.03^{+0.05}_{-0.06} & 0.02^{+0.07}_{-0.08} & -0.01^{+0.08}_{-0.08} \\
0.10^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.06} & -0.16^{+0.05}_{-0.05} & 0.03^{+0.05}_{-0.06} & 1 & 0.05^{+0.08}_{-0.08} & -0.01^{+0.07}_{-0.07} \\
0.00^{+0.08}_{-0.07} & -0.00^{+0.07}_{-0.08} & -0.03^{+0.08}_{-0.08} & 0.02^{+0.07}_{-0.08} & 0.05^{+0.08}_{-0.08} & 1 & -0.00^{+0.07}_{-0.08} \\
-0.01^{+0.08}_{-0.08} & -0.00^{+0.08}_{-0.08} & 0.01^{+0.08}_{-0.08} & -0.01^{+0.08}_{-0.08} & -0.01^{+0.07}_{-0.07} & -0.00^{+0.07}_{-0.08} & 1 \\
\end{array}
\end{pmatrix}.
\end{multline}
Their values for all Monte Carlo links for all supernovae are shown in Figure~\ref{perobject2:fig}.
There is a core concentration in the  parameter-space, with around ten objects that occupy its outskirts.
Many outliers appear in the red tail of $E_{\gamma^0}(B-V)$, as would be expected for the (infrequent) selection of supernovae
heavily extinguished by host-galaxy dust.
Outliers  are also clearly distinguishable in  $EW_{Ca}$--$\lambda_{Si}$ space.

@@

\subsection{Physical Interpretation of the Intrinsic Color Parameter}

Based on its direction, a straightforward explanation for the third color term $\pmb{\delta} D$ is that it describes the absorption due to some form of dust
that does not exactly obey the  \citetalias{1999PASP..111...63F} model.
In one scenario,
supernova light can encounter two
dusts, first a 
circumstellar dust that produces relatively low visual-to-selective extinction
\citep{2008ApJ...686L.103G} and then
an interstellar form  that resembles the \citetalias{1999PASP..111...63F} model.
To first order $\pmb{\delta}$ behaves like standard dust;   $>98$\% of its length projects onto the expectation of  \citetalias{1999PASP..111...63F}
with constant $R^F$.
The departure from   \citetalias{1999PASP..111...63F}
can be seen qualitatively; the decrease in magnitude change toward redder wavelengths, seen in Figure~\ref{deltaratio:fig}, matches the 
expectation for the circumstellar form of dust.
However, we
lack the signal-to-noise to show consistency with this explanation.
The best-fit elements $\delta_X$ do not all share the same sign, as would have been expected for dust.
Confining ourselves to the 98\% of   $\pmb{\delta}$ that lies on the dust plane, associating 
$D$ with $A_V$ would imply an unphysical $R^F=0.5$.

An alternative scenario is that supernova light encounters one dust with a degree of freedom beyond that modeled by \citetalias{1999PASP..111...63F}.
The $\pmb{\delta} D$ term represents an intrinsic dust parameter that is independent of $R^F$.
The extra degree-of-freedom could be associated with the skewness in the distribution of dust grains, as an increase of smaller dust grains
can flatten extinction in the red and enhanced steepness in the UV relative to standard dust models  \citep[Figure~3 in][]{2015ApJ...807L..26G,
2017ApJ...836...13H}.
Unfortunately, our signal-to-noise for $\pmb{\delta}$ is insufficient to test these physically-motivated
dust models.
\color{red}
\color{black}

>>>>

Each supernova is described by its parameters
 $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, $E_{\gamma^1}(B-V)$,  $A_{\delta U}=\delta_U D$.
 Their Pearson correlation coefficients are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V),  A_{\delta U}) =\\
\begin{pmatrix}
\begin{array}{rrrrrrr}
1 & 0.01^{+0.06}_{-0.06} & -0.07^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.07} & 0.10^{+0.07}_{-0.06} & 0.00^{+0.08}_{-0.07} & -0.01^{+0.08}_{-0.08} \\
0.01^{+0.06}_{-0.06} & 1 & 0.10^{+0.04}_{-0.04} & -0.26^{+0.03}_{-0.03} & -0.04^{+0.06}_{-0.06} & -0.00^{+0.07}_{-0.08} & -0.00^{+0.08}_{-0.08} \\
-0.07^{+0.07}_{-0.06} & 0.10^{+0.04}_{-0.04} & 1 & -0.14^{+0.03}_{-0.03} & -0.16^{+0.05}_{-0.05} & -0.03^{+0.08}_{-0.08} & 0.01^{+0.08}_{-0.08} \\
-0.04^{+0.06}_{-0.07} & -0.26^{+0.03}_{-0.03} & -0.14^{+0.03}_{-0.03} & 1 & 0.03^{+0.05}_{-0.06} & 0.02^{+0.07}_{-0.08} & -0.01^{+0.08}_{-0.08} \\
0.10^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.06} & -0.16^{+0.05}_{-0.05} & 0.03^{+0.05}_{-0.06} & 1 & 0.05^{+0.08}_{-0.08} & -0.01^{+0.07}_{-0.07} \\
0.00^{+0.08}_{-0.07} & -0.00^{+0.07}_{-0.08} & -0.03^{+0.08}_{-0.08} & 0.02^{+0.07}_{-0.08} & 0.05^{+0.08}_{-0.08} & 1 & -0.00^{+0.07}_{-0.08} \\
-0.01^{+0.08}_{-0.08} & -0.00^{+0.08}_{-0.08} & 0.01^{+0.08}_{-0.08} & -0.01^{+0.08}_{-0.08} & -0.01^{+0.07}_{-0.07} & -0.00^{+0.07}_{-0.08} & 1 \\
\end{array}
\end{pmatrix}.
\label{corr2p:eqn}
\end{multline}
Their values for all Monte Carlo links for all supernovae are shown in Figure~\ref{perobject2:fig}.
There is a core concentration in the  parameter-space, with around ten objects that occupy its outskirts.
Many outliers appear in the red tail of $E_{\gamma^0}(B-V)$, as would be expected for the (infrequent) selection of supernovae
heavily extinguished by host-galaxy dust.
Outliers  are also clearly distinguishable in  $EW_{Ca}$--$\lambda_{Si}$ space.

@@

\subsection{Physical Interpretation of the Intrinsic Color Parameter}
\color{red}
Supernova colors are expected to be affected by interstellar Milky Way-like dust (e.g.\ as modeled by  \citetalias{1999PASP..111...63F})
and by ``intrinsic'' color variations from the supernova itself and its surrounding circumstellar dust. Physical modeling
shows that the complexity of interstellar dust shouldn't be expected to be captured by a single parameter \citep{2015ApJ...807L..26G,
2017ApJ...836...13H}. Supernova models \citep[e.g.][]{2007ApJ...662..487W,2013MNRAS.436..333S,2014MNRAS.441..532D,2015MNRAS.454.2549B,
2017ApJ...846...58H} and circumstellar dust \citep{2008ApJ...686L.103G} also predict color variations not  captured
by  \citetalias{1999PASP..111...63F}.
The $D$ parameter
can be ascribed to either an increased accuracy in the description of interstellar dust or to a separate intrinsic mechanism.

Evidence for the intrinsic origin of $\pmb{\delta}D$ is seen in the relation between observed magnitudes corrected for intrinsic magnitude correlations with
spectral features (but not for the color parameters)
and observed  color.  Figure~\ref{speccorr:fig} shows such a plot for corrected observed $B_o$ magnitudes
as a function observed $B_o-V_o$, in correspondence with Figure~3 of  \citet{2017ApJ...842...93M}.  Supernovae with extreme blue colors exhibit a shallower slope than those
with extreme red colors.   \citet{2017ApJ...842...93M} have demonstrated that this is the expected behavior for
dust with constant $\beta$ (the SALT2 equivalent of $R$) with an exponential extinction distribution and an intrinsic source of color variablilty
with a smaller value of $\beta$.  Without interstellar dust, supernovae are bluer and take the intrinsic $\beta$ slope for the color-magnitude
relation.  When the intrinsic contribution to color is dwarfed by dust, supernovae are bluer and take the larger dust $\beta$ slope.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output25/speccorr_color.pdf}
      \caption{
      \color{red}Observed $B$-band magnitude corrected for intrinsic magnitude correlations with spectral features as a function of observed color $B_o-V_o$.
      The two solid lines are the linear regressions for points with abscissa values less than and greater than zero.
	\color{black}
   \label{speccorr:fig}}
\end{figure}

In contrast, if the colors were due to a single physical dust described by three parameters,
the observed correlation between the color-magnitude slope and observed color seen in Figure~\ref{speccorr:fig}
would manifest itself in correlation between the $D$,
$k_0$ and $k_1$ parameters.  To the contrary,
the correlations between these parameters  are consistent with zero as given  in Eqn.~\ref{corr2p:eqn}.

If  $\delta D$  is intrinsic, it can be attributed to an intrinsic supernova property  or a dust not described by $\gamma k$.
To first order $\pmb{\delta}$ behaves like standard dust;   $>98$\% of its length projects onto the expectation of  \citetalias{1999PASP..111...63F}
with constant $R^F$.
Therefore, a straightforward explanation is that it describes the absorption due to some form of dust
that does not exactly obey the  \citetalias{1999PASP..111...63F} model.
The departure from   \citetalias{1999PASP..111...63F}
can be seen qualitatively; the decrease in magnitude change toward redder wavelengths, seen in Figure~\ref{deltaratio:fig}, matches the 
expectation for the circumstellar form of dust.
However, we
lack the signal-to-noise to show consistency with this explanation.
The best-fit elements $\delta_X$ do not all share the same sign, as would have been expected for dust.
Confining ourselves to the 98\% of   $\pmb{\delta}$ that lies on the dust plane, associating 
$D$ with $A_V$ would imply an unphysical $R^F=0.5$.
We do not have conclusive evidence that  $\delta$ is being a property of a different form of dust, rather
than representing intrinsic supernova color variation.
\color{black}
